\texttt{Pythran} -- C++ for Snakes\\\emph{\small Static Compiler for High Performance}

### 2014-05-04 doc/papers/pydata2014/talk.tex
Update, thanks Serge

\date{PyData - May 4th 2014}

@@

\begin{frame}
  \maketitle

  \begin{center}
    \includegraphics[height=3em]{logo}
  \end{center}
  \vfill
  \hfill
  \scriptsize{Get the slides: http://goo.gl/6dgra0}
\end{frame}

@@

\begin{frame}[fragile]
    \frametitle{Static Compilation}
    \begin{block}{Buys time for many time-consuming analyse}
        Points-to, used-def, variable scope, memory effects, function purity\dots
    \end{block}
    \begin{block}{Unleashes powerful C++ optimizations}
        Lazy Evaluation, Map Parallelizations, Constant Folding
    \end{block}
    \begin{block}{Requires static type inference}
        \begin{lstlisting}[lang=python]
        #pythran export foo(int list, float)
        \end{lstlisting}
        \centering Only annotate exported functions!
    \end{block}
\end{frame}

@@

\begin{frame}[fragile]
    \frametitle{Explicit Parallelization}
            \begin{lstlisting}[basicstyle=\ttfamily\scriptsize]
def hyantes(xmin, ymin, xmax, ymax, step, range_, range_x, range_y, t):
  pt = [[0]*range_y for _ in range(range_x)]
  "omp parallel for"
  for i in xrange(range_x):
    for j in xrange(range_y):
      s = 0
      for k in t:
        tmp = 6368.* math.acos(math.cos(xmin+step*i)*math.cos( k[0] ) *
                               math.cos((ymin+step*j)-k[1]) +
                               math.sin(xmin+step*i)*math.sin(k[0]))
        if tmp < range_:
          s+=k[2] / (1+tmp)
      pt[i][j] = s
  return pt
\end{lstlisting}
\vfill
    \begin{tabular}{|l|r|r|r|}
        \hline
     Tool    &  CPython    &   Pythran     & OpenMP   \\
    \hline
     Timing  &  639.0ms    &   44.8ms       &      11.2ms       \\
    \hline
     Speedup &  $\times$1         &    $\times$14.2      &    $\times$57.7   \\
    \hline
\end{tabular}
\end{frame}

@@

\begin{lstlisting}
def run_julia(cr, ci, N, bound, lim, cutoff):
    julia = np.empty((N, N), np.uint32)
    grid_x = np.linspace(-bound, bound, N)
    t0 = time()
    "omp parallel for"
    for i, x in enumerate(grid_x):
        for j, y in enumerate(grid_x):
            julia[i,j] = kernel(x, y, cr, ci, lim, cutoff)
    return julia, time() - t0
\end{lstlisting}
From Scipy2013 Cython Tutorial.
\vfill
    \begin{tabular}{|l|r|r|r|r|}
        \hline
        Tool    &  CPython    &  Cython  &  Pythran & +OpenMP   \\
        \hline
        Timing  &  3630ms    &   4.3ms  &   3.71ms  & 1.52ms  \\
        \hline
        Speedup & $\times$1 &  $\times$837 &  $\times$970 & $\times$2368   \\
        \hline
    \end{tabular}

@@

\end{lstlisting}
\end{minipage}
\begin{minipage}{0.49\textwidth} \begin{lstlisting}[language=Python,frame=none,backgroundcolor=\color{white},basicstyle=\ttfamily\tiny]
#pythran export stddev(float64 list list)
def stddev(partitions):
    ddof=1
    final = 0.0
    for part in partitions:
        m = total(part) / len(part)
        # Find the mean of the entire grup.
        gtotal = total([total(p) for p in partitions])
        glength = total([len(p) for p in partitions])
        g = gtotal / glength
        adj = ((2 * total(part) * (m - g)) + ((g ** 2 - m ** 2) * len(part)))
        final += varsum(part) + adj
    return math.sqrt(final / (glength - ddof))
\end{lstlisting}
\end{minipage}
\vfill
\begin{tabular}{|l|r|r|r|r|r|r|}
  \hline
  Version  &  Awful     &  Less Awful  &  OK & \textit{Differently OK} & \textit{OK} & Numpy   \\
  \hline
  CPython  & 127s       &   150ms  &   54.3ms  & 53.8ms    & 47.6ms & 8.2ms  \\
  \hline
  Pythran  & 1.38s      &   4.7ms  &   4.8ms   &  5.8ms    &  4.7ms & 1.7ms \\
        \hline
    \end{tabular}

@@

\begin{frame}{Conclusions}
    \hfill\tiny{\url{http://pythonhosted.org/pythran/}}
    \normalsize
    \begin{center}
        \vfill
        Compiling Python means more than typing and translating\\
        \vfill
        \uncover<2-> {
        What next: 
        \begin{itemize}
          \item<+-> Release the PyData version (ooops we're late)
          \item<+-> User module import (pull request already issued)
          \item<+-> Set-up an daily performance regression test bot (ongoing work with codespeed)
          \item<+-> Re-enable vectorization through boost::simd and nt2 (soon)
          \item<+-> More Numpy support, start looking into Scipy
          \item<+-> Polyhedral transformations (for my next life...)
        \end{itemize}
        }
        \vfill
        % \uncover<3->{
        %     Need a compiler guy / HPC Engineer? I am looking for a job!\\
        % }
        \vfill
    \end{center}
\end{frame}

>>>>

\date{\small{PyData - May 4th 2014}}

@@

\begin{frame}
  \maketitle
\vspace{-0.5cm}
  \begin{center}
    \includegraphics[height=6em]{logo}
  \end{center}
  \vfill
  \hfill
  \scriptsize{Get the slides: http://goo.gl/6dgra0}
\end{frame}

@@

\begin{frame}[fragile]
    \frametitle{Static Compilation}
    \begin{block}{Buys time for many time-consuming analyses}
        Points-to, used-def, variable scope, memory effects, function purity\dots
    \end{block}
    \begin{block}{Unleashes powerful C++ optimizations}
        Lazy Evaluation, Map Parallelizations, Constant Folding
    \end{block}
    \begin{block}{Requires static type inference}
        \begin{lstlisting}[lang=python]
        #pythran export foo(int list, float)
        \end{lstlisting}
        \centering Only annotate exported functions!
    \end{block}
\end{frame}

@@

\begin{frame}[fragile]
    \frametitle{Explicit Parallelization}
            \begin{lstlisting}[basicstyle=\ttfamily\scriptsize]
def hyantes(xmin, ymin, xmax, ymax, step, range_, range_x, range_y, t):
  pt = [[0]*range_y for _ in range(range_x)]
  "#omp parallel for"
  for i in xrange(range_x):
    for j in xrange(range_y):
      s = 0
      for k in t:
        tmp = 6368.* math.acos(math.cos(xmin+step*i)*math.cos( k[0] ) *
                               math.cos((ymin+step*j)-k[1]) +
                               math.sin(xmin+step*i)*math.sin(k[0]))
        if tmp < range_:
          s+=k[2] / (1+tmp)
      pt[i][j] = s
  return pt
\end{lstlisting}
\vfill
    \begin{tabular}{|l|r|r|r|}
        \hline
     Tool    &  CPython    &   Pythran     & OpenMP   \\
    \hline
     Timing  &  639.0ms    &   44.8ms       &      11.2ms       \\
    \hline
     Speedup &  $\times$1         &    $\times$14.2      &    $\times$57.7   \\
    \hline
\end{tabular}
\end{frame}

@@

\begin{lstlisting}
def run_julia(cr, ci, N, bound, lim, cutoff):
    julia = np.empty((N, N), np.uint32)
    grid_x = np.linspace(-bound, bound, N)
    t0 = time()
    "#omp parallel for"
    for i, x in enumerate(grid_x):
        for j, y in enumerate(grid_x):
            julia[i,j] = kernel(x, y, cr, ci, lim, cutoff)
    return julia, time() - t0
\end{lstlisting}
From Scipy2013 Cython Tutorial.
\vfill
    \begin{tabular}{|l|r|r|r|r|}
        \hline
        Tool    &  CPython    &  Cython  &  Pythran & +OpenMP   \\
        \hline
        Timing  &  3630ms    &   4.3ms  &   3.71ms  & 1.52ms  \\
        \hline
        Speedup & $\times$1 &  $\times$837 &  $\times$970 & $\times$2368   \\
        \hline
    \end{tabular}

@@

\end{lstlisting}
\end{minipage}
\begin{minipage}{0.49\textwidth} \begin{lstlisting}[language=Python,frame=none,backgroundcolor=\color{white},basicstyle=\ttfamily\tiny]
#pythran export stddev(float64 list list)
def stddev(partitions):
    ddof=1
    final = 0.0
    for part in partitions:
        m = total(part) / len(part)
        # Find the mean of the entire grup.
        gtotal = total([total(p) for p in partitions])
        glength = total([len(p) for p in partitions])
        g = gtotal / glength
        adj = ((2 * total(part) * (m - g)) + ((g ** 2 - m ** 2) * len(part)))
        final += varsum(part) + adj
    return math.sqrt(final / (glength - ddof))
\end{lstlisting}
\end{minipage}
\vfill
\begin{tabular}{|l|r|r|r|r|r|r|}
  \hline
  Version  &  Awful     &  Less Awful  &  OK & \textit{Differently OK} & \textit{OK} & Numpy   \\
  \hline
  CPython  & 127s       &   150ms  &   54.3ms  & 53.8ms    & 47.6ms & 8.2ms  \\
  \hline

  Pythran  & \uncover<2-> {1.38s}      &   \uncover<3-> {4.7ms}  &   \uncover<4-> {4.8ms}   &  \uncover<5-> {5.8ms}    &  \uncover<6-> {4.7ms} & \uncover<7-> {1.7ms} \\
        \hline
    \end{tabular}

@@

\begin{frame}{Conclusions}
    \hfill\tiny{\url{http://pythonhosted.org/pythran/}}
    \normalsize
    \begin{center}
        \vfill
        Compiling Python means more than typing and translating\\
        \vfill
        \uncover<2-> {
        What next: 
        \begin{itemize}
          \item<3-> Release the PyData version \scriptsize{(ooops we're late)}
          \item<4-> User module import \scriptsize{(pull request already issued)}
          \item<5-> Set-up an daily performance regression test bot \scriptsize{(ongoing work with codespeed)}
          \item<6-> Re-enable vectorization through boost::simd and nt2 \scriptsize{(soon)}
          \item<7-> More Numpy support, start looking into Scipy
          \item<8-> Polyhedral transformations \scriptsize{(in another life...)}
        \end{itemize}
        }
        \vfill
        % \uncover<3->{
        %     Need a compiler guy / HPC Engineer? I am looking for a job!\\
        % }
        \vfill
    \end{center}
\end{frame}
